Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-08-23 doc/RMD/MLG-distribution.Rmd
change MLG to MLH

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
     # title = "Shared MLGs across regions",
      fill = "Number of\nMLGs",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper MLG"#,
      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
    ))
}

@@

```{r}
g <- graph_from_adjacency_matrix(matrix(0, 2, 2))
V(g)$size <- c(56, 56 - 32)
V(g)$label <- c("Number of MLGs in Region", "Number of private MLGs")
lay <- as.data.frame(matrix(0, 2, 2, dimnames = list(NULL, c("x", "y"))))
glegend <- create_layout(g, "manual", node.positions = as.data.frame(lay)) %>% ggraph() + 
  geom_node_circle(aes(r = size, fill = size)) + 
  coord_fixed() + 
  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/56, end = 1, guide = "none") + 
  ggrepel::geom_text_repel(aes(label = label), 
                           x = c(20, 0) + 25, 
                           y = c(35, 0), 
                           nudge_x = 350, 
                           segment.size = 1, 
                           arrow = arrow(length = unit(0.15, "native")), 
                           family = "Helvetica") + 
  theme_void() +
  # theme(plot.background = element_rect(color = "black", fill = "grey98")) +
  xlim(c(-60, 480))

>>>>

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
     # title = "Shared MLGs across regions",
      fill = "Number of\nMLHs",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper MLH"#,
      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
    ))
}

@@

```{r}
g <- graph_from_adjacency_matrix(matrix(0, 2, 2))
V(g)$size <- c(56, 56 - 32)
V(g)$label <- c("Number of MLHs in Region", "Number of private MLHs")
lay <- as.data.frame(matrix(0, 2, 2, dimnames = list(NULL, c("x", "y"))))
glegend <- create_layout(g, "manual", node.positions = as.data.frame(lay)) %>% ggraph() + 
  geom_node_circle(aes(r = size, fill = size)) + 
  coord_fixed() + 
  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/56, end = 1, guide = "none") + 
  ggrepel::geom_text_repel(aes(label = label), 
                           x = c(20, 0) + 25, 
                           y = c(35, 0), 
                           nudge_x = 350, 
                           segment.size = 1, 
                           arrow = arrow(length = unit(0.15, "native")), 
                           family = "Helvetica") + 
  theme_void() +
  # theme(plot.background = element_rect(color = "black", fill = "grey98")) +
  xlim(c(-60, 480))

### 2017-08-23 doc/RMD/mlg-mcg.Rmd
change MLG to MLH

```{r barplots, fig.width = 10, fig.height = 5}
mcg_graph <- mcgs %>% 
  filter(Samples > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.5, alpha = 0, lty = 2) +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  # scale_alpha(breaks = c(0.5, 0.75, 1)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  # coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.major.y = element_line(color = "grey50")) +
  theme(panel.grid.minor.y = element_line(color = "grey70")) +
  theme(legend.box.just = "top") +
  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
  theme(legend.box = "horizontal") +
  theme(legend.position = c(0.77, 0.75)) +
  theme(aspect.ratio = 1/2) +
  labs(list(
    alpha = "MLG\nEvenness"
  ))

@@

in_reach_summary <- in_reach %>% 
  group_by(N) %>%
  summarize(E = diversity_stats(as.integer(table(unlist(MCG))))["E.5"], 
            MCG = n_distinct(unlist(MCG)), 
            MLG = n())
knitr::kable(in_reach_summary)
in_reach_plot <- ggplot(in_reach_summary, aes(x = N, y = MLG, fill = E)) +
  geom_segment(aes(yend = 0, xend = N)) +
  geom_point(aes(size = MCG), pch = 21) +
  # ggrepel::geom_label_repel(aes(label = MCG)) +
  viridis::scale_fill_viridis(option = "D") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/1.75) +
  theme(axis.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0, 0.1), limits = c(0, 25)) +
  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-2, 60)) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(panel.border = element_blank()) +
  labs(list(fill = expression(E[5]), 
            size = "MCG"
            # caption = "size and fill represent\nthe number and distribution\nof MCG"
            )) +
  xlab("Potential Heterothallic Pairings") +
  ylab("Multilocus Genotypes")
in_reach_plot
if (!interactive()){
  save(in_reach, in_reach_summary, in_reach_plot, file = file.path(PROJHOME, "data/heterothallic-pairings.rda"))
}
```

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% (mcgs %>% arrange(desc(Samples)) %>% slice(1:8) %>% pull(MCG)))
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
# set.seed(2017-05-08)
set.seed(2017-08-1)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
levels(top5lay$type) <- c("Multilocus Genotype", "Mycelial Compatibility\nGroup")
t5g <- ggraph(top5lay) +
  geom_node_circle(aes(r = size, fill = type)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = type, 
                     size = size/10),
                 show.legend = FALSE) +
  coord_fixed() +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_fill_manual(values = c("black", "white")) +
  scale_color_manual(values = c("white", "black")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  # theme(legend.position = c(0.85, 0.35), legend.direction = "vertical") +
  theme(legend.position = c(0.5, 0), legend.direction = "horizontal", legend.box = "horizontal") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(legend.spacing.x = unit(0, "lines")) +
  # theme(legend.box.background = element_rect(color = "black")) +
  theme(legend.text = element_text(size = 10)) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
t5g$labels$edge_width <- "shared isolates"
t5g$labels$fill <- ""
t5g

@@

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Genotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
if (!interactive())
  ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
```

>>>>

```{r barplots, fig.width = 10, fig.height = 5}
mcg_graph <- mcgs %>% 
  filter(Samples > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  rename(Type = type) %>%
  mutate(Type = ifelse(Type == "MLGs", "MLHs", Type)) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLHs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.5, alpha = 0, lty = 2) +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  # scale_alpha(breaks = c(0.5, 0.75, 1)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  # coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.major.y = element_line(color = "grey50")) +
  theme(panel.grid.minor.y = element_line(color = "grey70")) +
  theme(legend.box.just = "top") +
  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
  theme(legend.box = "horizontal") +
  theme(legend.position = c(0.77, 0.75)) +
  theme(aspect.ratio = 1/2) +
  labs(list(
    alpha = "MLH\nEvenness"
  ))

@@

in_reach_summary <- in_reach %>% 
  group_by(N) %>%
  summarize(E = diversity_stats(as.integer(table(unlist(MCG))))["E.5"], 
            MCG = n_distinct(unlist(MCG)), 
            MLG = n())
knitr::kable(in_reach_summary)
in_reach_plot <- ggplot(in_reach_summary, aes(x = N, y = MLG, fill = E)) +
  geom_segment(aes(yend = 0, xend = N)) +
  geom_point(aes(size = MCG), pch = 21) +
  # ggrepel::geom_label_repel(aes(label = MCG)) +
  viridis::scale_fill_viridis(option = "D") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/1.75) +
  theme(axis.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0, 0.1), limits = c(0, 25)) +
  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-2, 60)) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(panel.border = element_blank()) +
  labs(list(fill = expression(E[5]), 
            size = "MCG"
            # caption = "size and fill represent\nthe number and distribution\nof MCG"
            )) +
  xlab("Potential Heterothallic Pairings") +
  ylab("Multilocus Haplotypes")
in_reach_plot
if (!interactive()){
  save(in_reach, in_reach_summary, in_reach_plot, file = file.path(PROJHOME, "data/heterothallic-pairings.rda"))
}
```

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% (mcgs %>% arrange(desc(Samples)) %>% slice(1:8) %>% pull(MCG)))
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
# set.seed(2017-05-08)
set.seed(2017-08-1)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
levels(top5lay$type) <- c("Multilocus Haplotype", "Mycelial Compatibility\nGroup")
t5g <- ggraph(top5lay) +
  geom_node_circle(aes(r = size, fill = type)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = type, 
                     size = size/10),
                 show.legend = FALSE) +
  coord_fixed() +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_fill_manual(values = c("black", "white")) +
  scale_color_manual(values = c("white", "black")) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  # theme(legend.position = c(0.85, 0.35), legend.direction = "vertical") +
  theme(legend.position = c(0.5, 0), legend.direction = "horizontal", legend.box = "horizontal") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(legend.spacing.x = unit(0, "lines")) +
  # theme(legend.box.background = element_rect(color = "black")) +
  theme(legend.text = element_text(size = 10)) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
t5g$labels$edge_width <- "shared isolates"
t5g$labels$fill <- ""
t5g

@@

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Haplotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
if (!interactive())
  ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
```
